Loop-Aware Optimizations in PyPy's Tracing JIT

### 2011-06-16 talk/iwtc11/paper.tex
fix url, use different bibliography style
always optimize original trace

The work described in this paper was done in the context of the PyPy
project\footnote{\url{http://pypy.org}}. PyPy is a framework for implementing
dynamic languages efficiently \cite{armin_rigo_pypys_2006}. When implementing a
language with PyPy, one writes an interpreter for the language in RPython
\cite{davide_ancona_rpython:_2007}. RPython (``Restricted Python``) is a subset
of Python chosen in such a way that it can be efficiently translated to a
C-based VM by performing type inference.

@@

$L_1$($p_{0}$, $p_{5}$, $i_3$, $i_4$):
# inside f: y = y.add(step)
    # inside BoxedInteger.add
        # inside BoxedInteger.add__int
        $i_{8}$ = $i_{4}+i_{3}$
        $p_{9}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{9}$, intval, $i_{8}$)
jump($L_1$, $p_{0}$, $p_{9}$, $i_3$, $i_8$)
\end{lstlisting}

@@

\subsection{Conclusions}
In this paper we have studied loop invariant code motion during trace
compilation. We claim that loop peeling is a very convenient solution
here since it fits well with other trace optimizations. By peeling of
the first iteration and optimizing the resulting two iteration trace
as a single trace, several standard optimizations can be 
used unchanged. The only interaction needed between the loop peeling
and the other 
optimizations is during the constructing of the jump arguments
connecting the peeled of iteration (the preamble) with the loop body. This
approach 
improves the effect of standard optimizations such as redundant guard removal, heap
caching, common subexpression elimination and allocation removals. The
most prominent effect is that they all become loop
invariant code motion optimizations.

XXX: is ``loop body'' or ``peeled loop'' the preferable term?

@@

\bibliographystyle{abbrvnat}
\bibliography{paper}

>>>>

The work described in this paper was done in the context of the PyPy
project\footnote{\texttt{http://pypy.org}}. PyPy is a framework for implementing
dynamic languages efficiently \cite{armin_rigo_pypys_2006}. When implementing a
language with PyPy, one writes an interpreter for the language in RPython
\cite{davide_ancona_rpython:_2007}. RPython (``Restricted Python``) is a subset
of Python chosen in such a way that it can be efficiently translated to a
C-based VM by performing type inference.

@@

$L_1$($p_{0}$, $p_{5}$, $i_3$, $i_4$):
# inside f: y = y.add(step)
guard_class($p_{5}$, BoxedInteger)
    # inside BoxedInteger.add
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{8}$ = $i_{4}+i_{3}$
        $p_{9}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{9}$, intval, $i_{8}$)
jump($L_1$, $p_{0}$, $p_{9}$, $i_3$, $i_8$)
\end{lstlisting}

@@

\subsection{Conclusions}
In this paper we have studied loop invariant code motion during trace
compilation. We claim that loop peeling is a very convenient solution
here since it fits well with other trace optimizations. By peeling of
the first iteration and optimizing the resulting two iteration trace
as a single trace, several standard optimizations can be 
used unchanged. The only interaction needed between the loop peeling
and the other 
optimizations is during the constructing of the jump arguments
connecting the peeled of iteration (the preamble) with the peeled loop. This
approach 
improves the effect of standard optimizations such as redundant guard removal, heap
caching, common subexpression elimination and allocation removals. The
most prominent effect is that they all become loop
invariant code motion optimizations.

@@

\bibliographystyle{abbrv}
\bibliography{paper}
